\documentclass[10pt]{article}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{enumitem}

\title{Implementation of Databases Sheet 1}
\author{Ilya Kulikov 351063, Alina Shigabutdinova, Oleg Chernikov}

\begin{document}
  \maketitle
  \section*{Exercise 1.1}
  \begin{enumerate}[label=\arabic*.] 
      \item
        There are layers in the database system:
        \begin{itemize}
          \item Logical data structures: this layer contains auxiliary structures:
          external schema description and integrity rules. The goal of this layer is
          to translate and optimize queries. Interface with transaction programs is
          SQL, which operates with relations, views and tuples.
          \item Logical access structures: this layer contains auxiliary structures:
          access path data, internal schema description. The task of this layer is to
          manage cursor, sort components and dictionary. Interface with logical data
          structures is record oriented interface, which operates with records, sets,
          keys, access paths.
          \item Storage structure: this layer contains auxiliary structures: DBTT, FPA,
          page indexes etc. The task of this layer is to manage record and index in the
          database system. Interface with Logical access structures is Internal record
          interface, which operates with records, B* trees etc.
          \item Page assignment: this layer contains auxiliary structures: page and
          block tables. The task of this layer is to manage buffer and segments. The
          interface with Storage structure is System buffer interface, which operates
          with pages and segments.
          \item Memory assignment structures: this layer contains auxiliary structures:
          VTOC, extent tables, system catalogue. The task of this layer is to manage files
          and external memory. The interface with Page assignment is File interface,
          which operates with blocks and files. The interface with Physical volume is
          Device interface, which operates with tracks, cylinders, channels etc.
        \end{itemize}
      \item
        The following sequence matches the top-down architecture:\\
        (E) $\Rightarrow$ (B) $\Rightarrow$ (D) $\Rightarrow$ (A) $\Rightarrow$ (C)\\\\
      \item
      	\begin{enumerate}[label=\alph*.] 
          \item Data independence means that application or some external representation of
          the data is independent from internal storage.\\
          It is not definitely given in the slides, but Data Independence contains
          two parts: Logical Data Independence and Physical Data Independence. Logical Data
          is a metadata and data about database itself, i.e. it describes how the actual
          data stored and managed in the database. So Logical Data Indeendence is a kind of
          mechanism, which acts independently from actual data stored on the disk. For example,
          if we do some change on table format, it should not change the data on the disk.\\
          Physical data independence is the mechanism which allows to change physical data
          without impact the schema or some other logical data.\\
          \item Data independence is an important feature because it is not feasible to store
          the same data for each application several times and to provide shared data to
          several applications and to be sure, that data will be accessible and consistent.
          More over, with data independence it is possible to use logical data structures
          and descriptive queries.\\
          \item Data structures and smart modern interfaces hiding all low-level layers and working
        with filesystems and block devices.
        \end{enumerate}
  \end{enumerate}
  
  \section*{Exercise 1.2}
  \begin{enumerate}[label=\arabic*.]
  \item
 	SELECT e1."EmployeeId", e1."LastName", e1."FirstName", e1."Phone", e1."Email" 
	FROM "Employee" e1, "Employee" e2 
	WHERE e2."EmployeeId" =  e1."ReportsTo" AND e1."HireDate" > e2."HireDate"
  \item
    SELECT sum("Total") FROM 
	(SELECT iid, count(alid) c FROM (SELECT DISTINCT i."InvoiceId" iid, i."CustomerId" cid, al."ArtistId" alid FROM "InvoiceLine" il, "Invoice" i, "Track" t, "Album" al
	WHERE i."InvoiceId" = il."InvoiceId" AND t."TrackId" = il."TrackId" AND al."AlbumId" = t."AlbumId" ORDER BY i."InvoiceId") temp1 GROUP BY iid) temp2,
	"Invoice" i
WHERE
i."InvoiceId" = iid AND c > 1 GROUP BY "CustomerId"
  \item
    SELECT sum(i."Total"), sum(t."Bytes")
    FROM "Playlist" p, "PlaylistTrack" pt, "InvoiceLine" il, "Invoice" i, "Track" t
    WHERE p."Name" = 'Grunge' AND pt."PlaylistId" = p."PlaylistId" AND il."TrackId" = pt."TrackId" AND il."InvoiceId" = i."InvoiceId" AND t."TrackId" = pt."TrackId"
  \item
    SELECT p."Name" FROM "Playlist" p
    WHERE p."PlaylistId" NOT IN (SELECT DISTINCT pt."PlaylistId" ptid FROM "Track" t, "Album" al, "Artist" ar, "PlaylistTrack" pt
    WHERE (ar."Name" = 'Black Sabbath' OR ar."Name" = 'Chico Buarque') AND al."ArtistId" = ar."ArtistId" AND t."AlbumId" = al."AlbumId" AND pt."TrackId" = t."TrackId")
\end{enumerate}

\end{document}
